---
title: "Codebook.md"
author: "Vladimir Okhotnikov"
date: "19 December 2014"
output:
  html_document:
    keep_md: yes
---

The initial raw data was obtained from 

[https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip)

as the README file of the raw data describes the process of capturing it:

> The experiments have been carried out with a group of 30 volunteers within an 
> age bracket of 19-48 years. Each person performed six activities (WALKING, 
> WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a 
> smartphone (Samsung Galaxy S II) on the waist. Using its embedded 
> accelerometer and gyroscope, we captured 3-axial linear acceleration and 
> 3-axial angular velocity at a constant rate of 50Hz. The experiments have 
> been video-recorded to label the data manually. The obtained dataset has been 
> randomly partitioned into two sets, where 70% of the volunteers was selected 
> for generating the training data and 30% the test data. 

> The sensor signals (accelerometer and gyroscope) were pre-processed by 
> applying noise filters and then sampled in fixed-width sliding windows of 
> 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration 
> signal, which has gravitational and body motion components, was separated 
> using a Butterworth low-pass filter into body acceleration and gravity. The 
> gravitational force is assumed to have only low frequency components, 
> therefore a filter with 0.3 Hz cutoff frequency was used. From each window, 
> a vector of features was obtained by calculating variables from the time 
> and frequency domain.

The raw data also provided jerk and magnitude variables:

> Subsequently, the body linear acceleration and angular velocity were derived 
> in time to obtain Jerk signals... Also the magnitude of these 
> three-dimensional signals were calculated using the Euclidean norm 

The resulting tidy data includes only mean value and standard deviation pairs 
for each variable, summarized by activity and subject pair. Variable names
have been sanitized to avoid unsave characters (like parentheses) and obvious
typos (like BodyBody)

```{r, echo = FALSE}
d <- read.table("tidy/step5.txt", header = TRUE)
```

The `Activity` column describes the activity performed by the subject. It is a 
factor with `r length(levels(d$Activity))` levels: 
```{r, echo = FALSE}
levels(d$Activity)
```

The `Subject` columns identifies a particular subject performing the activity. 
It is an integer between 1 and 30 (inclusive).

The remaining columns represent mean and standard deviation values averaged by 
key pair of `Subject` and `Activity`. All values are normalized and bounded 
within [-1;1].

```{r, echo=FALSE}
options(width = 120)
deduceDescription <- function(name) {
    last <- function(x) { tail(x, n = 1) }
    stripLast <- function(x) { x[1:length(x)-1] }
    
    el = unlist(strsplit(name, "\\."))
    
    byAxis <- ""
    if (last(el) %in% c("X", "Y", "Z")) {
        l <- last(el)
        if (l == "X") {
            byAxis <- " by X axis"
        } else if (l == "Y") {
            byAxis <- " by Y axis"
        } else if (l == "Z") {
            byAxis <- " by Z axis"
        }
        el <- stripLast(el)
    }
    
    meanOrStd <- NA
    if (last(el) == "mean") {
        meanOrStd <- "Mean value"
        el <- stripLast(el)
    } else if (last(el) == "std") {
        meanOrStd <- "Standard deviation"
        el <- stripLast(el)
    }
    
    if (length(el) == 1) {
        knownVars <- as.data.frame(rbind(
            c("tBodyGyro", "body angular velocity in time domain"),
            c("fBodyGyro", "body angular velocity in frequency domain"),
            c("tBodyGyroJerk", "derived body angular velocity (Jerk) in time domain"),
            c("tBodyAccJerk", "derived body acceleration (Jerk) in time domain"),
            c("fBodyAcc", "body acceleraiton in frequency domain"),
            c("tGravityAcc", "gravity acceleration component in time domain"),
            c("tBodyAccMag", "magnitude of body acceleration component in time domain"),
            c("tBodyAccJerkMag", "magnitude of derived body acceleration (Jerk) in time domain"),
            c("tGravityAccMag", "magnitude of gravity acceleration component in time domain"),
            c("tBodyGyroMag", "magnitude of angular velocity component in time domain"),
            c("tBodyGyroJerkMag", "magnitude of derived angular velocity (Jerk) in time domain"),
            c("fBodyAccMag", "magnitude of body acceleration component in frequency domain"),
            c("fBodyAccJerkMag", "magnitude of derived body acceleration (Jerk) in frequency domain"),
            c("fBodyGyroMag", "magnitude of angular velocity component in frequency domain"),
            c("fBodyGyroJerkMag", "magnitude of derived angular velocity (Jerk) in frequency domain"),
            c("tBodyAcc", "body acceleration component in time domain"),
            c("fBodyAccJerk", "derived body acceleration (Jerk) in frequency domain")
            ))
        
        idx <- which(knownVars[,1] == el[1])
        if (length(idx) == 1) {
            paste(meanOrStd, " of ", knownVars[idx, 2], byAxis, sep = "")
        } else {
            NA
        }
    } else {
        "-"
    }
}

vars <- colnames(d)[3:ncol(d)]
desc <- sapply(vars, deduceDescription)
data.frame(variable = vars, description = desc, row.names = 3:ncol(d))
```